---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
```{r, echo=FALSE}
knitr::opts_chunk$set(echo=TRUE)
```

## Loading and preprocessing the data

Fork/Clone the repository and set the working directory to the cloned folder
```{r}
setwd("~/Coursera/Reproducible Research/PeerAssessment1/RepData_PeerAssessment1")
cn <- unz(description="activity.zip", filename="activity.csv")
act_data <- read.csv(cn)
act_data$date <- as.Date(act_data$date)
```

## What is mean total number of steps taken per day?

```{r}
library(dplyr)
```

```{r}
steps_day <- summarise(group_by(act_data, date), ttl_steps=sum(steps))
```

```{r, fig.height=3}
hist(steps_day$ttl_steps, breaks=25, main="Distribution of steps per day", xlab=NULL, ylab="Number of days")
```

```{r}
mn <- mean(steps_day$ttl_steps, na.rm=TRUE)
md <- median(steps_day$ttl_steps, na.rm=TRUE)
```

The mean and median of the total number of steps taken per day is `r format(mn)` and `r md` respectively.

## What is the average daily activity pattern?

```{r}
day_pat <- summarise(group_by(act_data, interval), avg_steps=mean(steps, na.rm=TRUE))
```

```{r, fig.height=4}
with(day_pat, plot(interval, avg_steps, type="l", main="Daily activity pattern", xlab="ID of 5-minute interval", ylab="Average number of steps"))
```

```{r}
mx_int <- day_pat$interval[which.max(day_pat$avg_steps)]
```

The 5-minute interval with ID = `r mx_int` contains the maximum number of steps on average across all the days in the dataset.

## Imputing missing values

```{r}
mv_num <- sum(is.na(act_data$steps))
```

The total number of missing values in the dataset is `r mv_num`.

```{r}
act_data_impd <- act_data
act_data_impd$steps[which(is.na(act_data_impd$steps))] <- day_pat$avg_steps[which(day_pat$interval %in% act_data_impd$interval[which(is.na(act_data_impd$steps))])]
```

```{r}
steps_day_impd <- summarise(group_by(act_data_impd, date), ttl_steps=sum(steps))
```

```{r, fig.height=3}
hist(steps_day_impd$ttl_steps, breaks=25, main="Distribution of steps per day after imputation", xlab=NULL, ylab="Number of days")
```

```{r}
mn_impd <- mean(steps_day_impd$ttl_steps, na.rm=TRUE)
md_impd <- median(steps_day_impd$ttl_steps, na.rm=TRUE)
```

The mean and median of the total number of steps taken per day is `r format(mn_impd)` and `r format(md_impd)` respectively.
Median has changed after imputation.

## Are there differences in activity patterns between weekdays and weekends?

```{r}
act_data_impd$wkday <- factor(weekdays(act_data_impd$date, abbreviate = TRUE) %in% c("бс", "Тё"), ordered=TRUE)
levels(act_data_impd$wkday) <- c("weekday", "weekend")
day_pat2 <- summarise(group_by(act_data_impd, interval, wkday), avg_steps=mean(steps, na.rm=TRUE))

library(ggplot2)
```

```{r, fig.height=8}

qplot(interval, avg_steps, data = day_pat2, facets = wkday ~ .,
          xlab="Interval",
          ylab="Number of steps", geom = "line")
```

```{r}
sessionInfo()
```